Predicting tree failure likelihood for utility risk mitigation via a novel convolutional neural network

### 2021-02-01 docs/manuscript/manuscript.tex
minor changes to results

\usepackage{fancyhdr}
\pagestyle{fancy}
%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
%\addtolength{\headwidth}{\marginparwidth}
%\fancyheadoffset{30pt}
%\fancyfootoffset{30pt}
\fancyhead[LO,RE]{\small  \it \nouppercase{\leftmark}}
\fancyhead[RO,LE]{\small Page \thepage}
\fancyfoot[RO,LE]{\small }% PR \num S-2015}
\fancyfoot[LO,RE]{\small }%\scshape MODL}
\cfoot{}
\renewcommand{\headrulewidth}{0.1pt}
\renewcommand{\footrulewidth}{0pt}
%\setlength\voffset{-0.25in}
%\setlength\textheight{648pt}

@@

Pruning, despite its high cost, has been found to be effective in reducing disruptions due to so-called 'preventable' contact incidents between trees and power lines.  For example, in Massachusetts, where tree failure was responsible for 40\% of preventable tree-caused outages, pruning was able to improve reliability by 20\% to 30\% \cite{simpson1996treecaused}, and those results were replicated in a study in Connecticut \cite{parent2019analysis}. The efficacy of pruning has also been shown in a study of two US Gulf Coast states that showed wind-induced power outage prediction models becoming less uncertain when pruning was included in the model \citet{nateghi2014power}.

@@

Individual tree risk assessment can be costly because of the time it requires. In some situations, a less time-consuming assessment may be justified to reduce costs, i.e.\ a ``Level 1'' assessment  \cite{e.thomassmiley2017best}. Studies have shown that trees with greater risk ratings were more likely to be detected from Level 1 risk assessments conducted in a moving vehicle in Rhode Island, USA \cite{rooney2005reliability} and Florida, USA \cite{koeser2016frequency} . The utility of Level 1 assessments in these areas suggests that artificial intelligence (AI) tools may be an effective way to reduce the cost of tree risk assessment.

@@

%Neural networks developed over the past several decades from the introduction of the single-layer perceptron.
The groundbreaking study of \citet{hubel1959receptive} showed that visual perception in cats was a result of the
activation or inhibition of groups of cells in the visual cortex known as ``receptive fields.''  Further, they attempted
to map the cortical architecture in cats and monkeys \cite{hubel1962receptive,hubel1965receptive,hubel1968receptive}.
Subsequent attempts were then made to model neural networks that could be trained to automatically recognize visual patterns with modest performance \cite{rosenblatt1962principles,kabrisky1966proposed,giebel1971feature,fukushima1975cognitron}. However, the breakthrough came with ``neocognitron'' \cite{fukushima1980neocognitron}, which
was a self-learning neural network for pattern recognition that was robust to changes in position and shape distortion, a problem that plagued earlier efforts, including ``cognitron'' \cite{fukushima1975cognitron} proposed a few years earlier.

@@

\begin{table}[h!]\small
  \centering
  \begin{tabular}{l l l l l l l }\toprule
    \bf Model & \multicolumn{3}{c}{\bf Training metrics} &\multicolumn{3}{c}{\bf Validation metrics}  \\\midrule
    & Error & Precision & Recall     & Error & Precision & Recall \\
    SafeTree & & & & & & \\
    GoogleNet (InceptionV3) & & & & & & \\
    ResNet50 & & & & & & \\
        VGGNet & & & & & & \\
    AlexNet & & & & & & \\\bottomrule
  \end{tabular}
  \caption{Comparing our model SafeTree with state-of-the-art CNN architectures trained on our data}
  \label{tab:comp}
\end{table}

@@

\section{Conclusion}

>>>>

% \usepackage{fancyhdr}
% \pagestyle{fancy}
%\addtolength{\headwidth}{\marginparsep} %these change header-rule width
%\addtolength{\headwidth}{\marginparwidth}
%\fancyheadoffset{30pt}
%\fancyfootoffset{30pt}
% \fancyhead[LO,RE]{\small  \it \nouppercase{\leftmark}}
% \fancyhead[RO,LE]{\small Page \thepage}
% \fancyfoot[RO,LE]{\small }% PR \num S-2015}
% \fancyfoot[LO,RE]{\small }%\scshape MODL}
% \cfoot{}
% \renewcommand{\headrulewidth}{0.1pt}
% \renewcommand{\footrulewidth}{0pt}
%\setlength\voffset{-0.25in}
%\setlength\textheight{648pt}

@@

Pruning, despite its high cost, has been found to be effective in reducing disruptions due to so-called ``preventable'' contact incidents between trees and power lines.  For example, in Massachusetts, where tree failure was responsible for 40\% of preventable tree-caused outages, pruning was able to improve reliability by 20\% to 30\% \cite{simpson1996treecaused}, and those results were replicated in a study in Connecticut \cite{parent2019analysis}. The efficacy of pruning has also been shown in a study of two US Gulf Coast states that showed wind-induced power outage prediction models becoming less uncertain when pruning was included in the model \citet{nateghi2014power}.

@@

Individual tree risk assessment can be costly because of the time it requires. In some situations, a less time-consuming assessment may be justified to reduce costs, i.e.\ a ``Level 1'' assessment  \cite{e.thomassmiley2017best}. Studies have shown that trees with greater risk ratings were more likely to be detected from Level 1 risk assessments conducted in a moving vehicle in Rhode Island, USA \cite{rooney2005reliability} and Florida, USA \cite{koeser2016frequency}. The utility of Level 1 assessments in these areas suggests that artificial intelligence (AI) tools may be an effective way to reduce the cost of tree risk assessment.

@@

%Neural networks developed over the past several decades from the introduction of the single-layer perceptron.
The groundbreaking study of \citet{hubel1959receptive} showed that visual perception in cats was a result of the
activation or inhibition of groups of cells in the visual cortex known as ``receptive fields.''  Further, they attempted
to map the cortical architecture in cats and monkeys \cite{hubel1962receptive,hubel1965receptive,hubel1968receptive}.
Subsequent attempts were then made to model neural networks that could be trained to automatically recognize visual patterns with modest performance \cite{rosenblatt1962principles,kabrisky1966proposed,giebel1971feature,fukushima1975cognitron}. However, the breakthrough came with the ``neocognitron'' \cite{fukushima1980neocognitron}, which
was a self-learning neural network for pattern recognition that was robust to changes in position and shape distortion, a problem that plagued earlier efforts, including ``cognitron'' \cite{fukushima1975cognitron} proposed a few years earlier.

@@

@@

\subsection{Model visualization and inference}

\subsection{Comparison with state-of-the-art architectures}
We compare the performance of our selected model with existing high-performance architectures. The results are summarized in \autoref{tab:comp}.

\begin{table}[h!]\small
  \centering
  \begin{tabular}{l l l l l l l }\toprule
    \bf Model & \multicolumn{3}{c}{\bf Training metrics} &\multicolumn{3}{c}{\bf Validation metrics}  \\\midrule
    & Error & Precision & Recall     & Error & Precision & Recall \\
    SafeTree & & & & & & \\
    GoogleNet (InceptionV3) & & & & & & \\
    ResNet50 & & & & & & \\
        VGGNet & & & & & & \\
    AlexNet & & & & & & \\\bottomrule
  \end{tabular}
  \caption{Comparing our model SafeTree with state-of-the-art CNN architectures trained on our data}
  \label{tab:comp}
\end{table}
\section{Conclusion}
